---
title: "Shiny_March_Madness_Data_Cleaning"
author: "Bailey Russo"
date: "4/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(shiny)
library(dplyr)
library(ggplot2)
library(tidyr)

relativePath = './March_Madness_Visualization/ncaam-march-mania-2021/'

path_out = '/Users/baileyrusso/Desktop/NYC Data Science Academy/Project_1/March_Madness_Visualization/ncaam-march-mania-2021//UserGeneratedDataFiles/'
```

```{r}

regSeasonAll <- read.csv(paste0(relativePath,'MDataFiles_Stage2/MRegularSeasonDetailedResults.csv'))

teams <- read.csv(paste0(relativePath,'MDataFiles_Stage1/MTeams.csv'))
teams = teams %>% select(.,TeamID, TeamName)
seasons <- c(2010:2021)

dataframes_all = c()
for (j in 1:length(seasons)) {
  
season = seasons[[j]]
regSeason = regSeasonAll %>% filter(.,Season==season)
dataframes_one_season = c()

for (i in 1:length(teams$TeamID)) {
  teamID = teams$TeamID[i]
  
  dataframes_one_season[[i]] = regSeason %>% filter(.,WTeamID == teamID| LTeamID == teamID) %>%  mutate(.,Score=ifelse(WTeamID==teamID,WScore,LScore),
              OppScore=ifelse(WTeamID==teamID,LScore,WScore),
              
              FGM=ifelse(WTeamID==teamID,WFGM,LFGM),
              OppFGM=ifelse(WTeamID==teamID,LFGM,WFGM),
              
              FGA=ifelse(WTeamID==teamID,WFGA,LFGA),
              OppFGA=ifelse(WTeamID==teamID,LFGA,WFGA),
              
              FGM3=ifelse(WTeamID==teamID,WFGM3,LFGM3),
              OppFGM3=ifelse(WTeamID==teamID,LFGM3,WFGM3),
              
              FGA3=ifelse(WTeamID==teamID,WFGA3,LFGA3),
              OppFGA3=ifelse(WTeamID==teamID,LFGA3,WFGA3),
              
              FTM=ifelse(WTeamID==teamID,WFTM,LFTM),
              OppFTM=ifelse(WTeamID==teamID,LFTM,WFTM),
              
              FTA=ifelse(WTeamID==teamID,WFTA,LFTA),
              OppFTA=ifelse(WTeamID==teamID,LFTA,WFTA),
              
              OR=ifelse(WTeamID==teamID,WOR,LOR),
              OppOR=ifelse(WTeamID==teamID,LOR,WOR),
              
              DR=ifelse(WTeamID==teamID,WDR,LDR),
              OppDR=ifelse(WTeamID==teamID,LDR,WDR),
              
              Ast=ifelse(WTeamID==teamID,WAst,LAst),
              OppAst=ifelse(WTeamID==teamID,LAst,WAst),
              
              TO=ifelse(WTeamID==teamID,WTO,LTO),
              OppTO=ifelse(WTeamID==teamID,LTO,WTO),
              
              Stl=ifelse(WTeamID==teamID,WStl,LStl),
              OppStl=ifelse(WTeamID==teamID,LStl,WStl),
              
              Blk=ifelse(WTeamID==teamID,WBlk,LBlk),
              OppBlk=ifelse(WTeamID==teamID,LBlk,WBlk),
              
              PF=ifelse(WTeamID==teamID,WPF,LPF),
              OppPF=ifelse(WTeamID==teamID,LPF,WPF),
              
              TeamID=ifelse(WTeamID==teamID,WTeamID,LTeamID),
              OppTeamID=ifelse(WTeamID==teamID,LTeamID,WTeamID),
              
              Outcome=ifelse(WTeamID==teamID, as.character('W'),as.character('L')),
              
              GameLoc=ifelse(WTeamID==teamID, as.character(WLoc), ifelse(WLoc==as.character('N'),as.character('N'),ifelse(WLoc==as.character('H'),as.character('A'),as.character('H'))))
              )
}

dataframes_all[[j]] = do.call("rbind", dataframes_one_season)

}

combined_df = do.call("rbind", dataframes_all)
combined_df = combined_df[, -c(3:34)]

combined_df = combined_df %>% mutate(.,FGPercentage=FGM/FGA) %>% mutate(.,OppFGPercentage=OppFGM/OppFGA) %>% mutate(.,ThreePercentage=FGM3/FGA3) %>% mutate(.,OppThreePercentage=OppFGM3/OppFGA3) %>% mutate(.,FTPercentage=FTM/FTA) %>% mutate(.,OppFTPercentage=OppFTM/OppFTA) %>% mutate(.,TSPercentage=Score/(2*(FGA + 0.44*FTA))) %>% mutate(.,OppTSPercentage=OppScore/(2*(OppFGA + 0.44*OppFTA))) %>% mutate(.,EffectiveFGPercentage=(FGM + 0.5*FGM3)/FGA) %>% mutate(.,OppEffectiveFGPercentage=(OppFGM + 0.5*OppFGM3)/OppFGA) %>% mutate(.,Possessions=(FGA - OR) + TO + (0.44*FTA)) %>% mutate(.,OppPossessions=(OppFGA - OppOR) + OppTO + (0.44*OppFTA)) %>% mutate(.,OE=(Score/Possessions)*100) %>% mutate(OppOE=(OppScore/OppPossessions)*100) %>% mutate(.,DE=(OppScore/OppPossessions)*100) %>% mutate(.,OppDE=(Score/Possessions)*100) %>% mutate(.,EM=OE-DE) %>% mutate(.,OppEM=DE-OE) %>% mutate(.,TurnoverPercentage=TO/(FGA +(0.44*FTA + TO))) %>% mutate(.,OppTurnoverPercentage=OppTO/(OppFGA +(0.44*OppFTA + OppTO))) %>% mutate(.,ORPercentage=OR/(OR + OppDR)) %>% mutate(.,OppORPercentage=OppOR/(OppOR + DR)) %>% mutate(.,DRPercentage=DR/(OppOR + DR)) %>% mutate(.,OppDRPercentage=OppDR/(OR + OppDR))
```

# Adjusting Stats
```{r}
############### Adjusting stats ##################

#Get the average statistic for each team at the season level
drop_cols = c('DayNum','Outcome','GameLoc','OppTeamID')
combined_df_team_avg = combined_df %>% select(-drop_cols) %>% group_by(.,Season,TeamID) %>% summarise(across(c("Score":"OppDRPercentage"),mean))

#Get the national average for each statistic
combined_df_avg = combined_df %>% select(-drop_cols) %>% group_by(.,Season,TeamID) %>% summarise(across(c("Score":"OppDRPercentage"),mean)) %>% group_by(.,Season) %>% summarise(across(c("Score":"OppDRPercentage"),mean))

#Rename all columns before merging
combined_df_team_avg = combined_df_team_avg %>% rename_all(~ paste0("AvgOpp_", names(combined_df_team_avg)))
combined_df_avg = combined_df_avg %>% rename_all(~ paste0("NationalAvg_",names(combined_df_avg)))

#Merge to get opponent seasonal averages and national averages
test = inner_join(combined_df, combined_df_team_avg, by=c("Season" = "AvgOpp_Season", "OppTeamID" = "AvgOpp_TeamID"))

test = inner_join(test, combined_df_avg,by=c("Season" = "NationalAvg_Season"))
test = inner_join(test, teams, by="TeamID")

############### Adjusting stats ##################

for (i in 3:58) {
  if (colnames(test)[i] == 'Outcome' | colnames(test)[i] == 'TeamID' | colnames(test)[i] == 'OppTeamID' | colnames(test)[i] == 'GameLoc') {
    print('Do not adjust this column.')
  } else {
  if (startsWith(colnames(test)[i], 'Opp')) {
    stat_name = as.name(paste0('Adjusted_',colnames(test)[i]))
    test = test %>% mutate(.,!!stat_name := test[,i] - ((get(paste0('AvgOpp_',substr(colnames(test)[i],4,nchar(colnames(test)[i]))))) - get(paste0('NationalAvg_',substr(colnames(test)[i],4,nchar(colnames(test)[i]))))))
  } else {
    stat_name = as.name(paste0('Adjusted_',colnames(test)[i]))
    test = test %>% mutate(.,!!stat_name := test[,i] - (get(paste0('AvgOpp_Opp',colnames(test)[i])) - get(paste0('NationalAvg_Opp',colnames(test)[i]))))
  }
  }
}

test = test[-c(59:162)]
test = test %>% select(.,-c('Adjusted_FTPercentage','Adjusted_OppFTPercentage'))

```

# Calculating Adjusted Tempo
```{r}
possession_stats_team = test %>% group_by(.,Season,TeamID) %>% summarise(PossessionsPerGame=mean(Possessions))

possession_stats_opp = test %>% group_by(.,Season,TeamID) %>% summarise(OppPossessionsPerGame=mean(Possessions))

test = inner_join(test,possession_stats_team,by=c('Season','TeamID'))
test = inner_join(test,possession_stats_opp,by=c('Season' = 'Season','OppTeamID' = 'TeamID'))

national_avg_possessions_per_game = test %>% group_by(.,Season,TeamID) %>% summarise(.,PossessionsPerGame=mean(Possessions)) %>% group_by(.,Season) %>% summarise(.,NationalAvg_PossessionsPerGame=mean(PossessionsPerGame))

test = inner_join(test, national_avg_possessions_per_game,by='Season')

test = test %>% mutate(.,ExpectedPossessions=NationalAvg_PossessionsPerGame + ( PossessionsPerGame - NationalAvg_PossessionsPerGame) + (OppPossessionsPerGame - NationalAvg_PossessionsPerGame)) %>% mutate(.,ActualPossessionsInGame=(Possessions + OppPossessions) / 2)

test = test %>% mutate(.,Difference=(ActualPossessionsInGame - ExpectedPossessions)) %>% mutate(.,SeasonPaceSum=PossessionsPerGame + OppPossessionsPerGame) %>% mutate(.,SeasonPacePercentage=PossessionsPerGame/SeasonPaceSum) %>% mutate(.,OppSeasonPacePercentage=OppPossessionsPerGame/SeasonPaceSum) %>% mutate(.,PaceShare=Difference*SeasonPacePercentage) %>% mutate(.,OppPaceShare=Difference*OppSeasonPacePercentage) %>% mutate(.,Adjusted_Tempo=PaceShare+PossessionsPerGame) %>% mutate(.,Adjusted_OppTempo=OppPaceShare+OppPossessionsPerGame) %>% mutate(.,check = NationalAvg_PossessionsPerGame + (Adjusted_Tempo-NationalAvg_PossessionsPerGame) + (Adjusted_OppTempo-NationalAvg_PossessionsPerGame))

team_averages = test %>% filter(.,Season=='2021') %>% group_by(.,Season,TeamName,TeamID) %>% summarise(.,across(c("Score":"Adjusted_OppTempo"),mean)) %>% select(.,Season,TeamID,TeamName,Adjusted_OE,Adjusted_DE,Adjusted_OppOE,Adjusted_OppDE,Adjusted_Tempo,Adjusted_OppTempo)

season_averages = test %>% filter(.,Season=='2021') %>% group_by(.,Season) %>% select(.,Season,TeamID,TeamName,Adjusted_OE,Adjusted_DE,Adjusted_OppOE,Adjusted_OppDE,Adjusted_Tempo,Adjusted_OppTempo) %>% summarise(across(c('Adjusted_OE':'Adjusted_OppTempo'),mean))

```

# Predicting Score Between Two Teams
```{r}
######### Predict Score For Two Teams ############
team_1 = team_averages %>% filter(.,TeamName=='Baylor')
team_2 = team_averages %>% filter(.,TeamName=='Wisconsin')

ExpectedOE_team1 = season_averages[,'Adjusted_OE'] + (team_1[,'Adjusted_OE'] - season_averages[,'Adjusted_OE']) + (team_2[,'Adjusted_DE'] - season_averages[,'Adjusted_OE'])

ExpectedOE_team2 = season_averages[,'Adjusted_OE'] + (team_2[,'Adjusted_OE'] - season_averages[,'Adjusted_OE']) + (team_1[,'Adjusted_DE'] - season_averages[,'Adjusted_OE'])

ExpectedTempo = season_averages[,'Adjusted_Tempo'] + (team_1[,'Adjusted_Tempo'] - season_averages[,'Adjusted_Tempo']) + (team_2[,'Adjusted_Tempo'] - season_averages[,'Adjusted_Tempo'])

Team1Score = ExpectedOE_team1/100 * ExpectedTempo 
Team2Score = ExpectedOE_team2/100 * ExpectedTempo
Team1Name = team_1[,'TeamName']
Team2Name = team_2[,'TeamName']
Team1Name <- data.frame(lapply(Team1Name, as.character), stringsAsFactors=FALSE)
Team2Name <- data.frame(lapply(Team2Name, as.character), stringsAsFactors=FALSE)

sprintf("%s : %s %s : %s", Team1Name, Team1Score, Team2Name, Team2Score)

test = test %>% select(-c('PossessionsPerGame':'OppPaceShare')) %>% select(.,-c("check"))

```

# Get dataset with non-adjusted stats
```{r}
teamWinPercentage = test %>% group_by(.,Season,TeamID) %>% summarise(.,WinPercentage=sum(Outcome=='W')/n())

RegSeasonData = test %>% select(.,-c('Adjusted_Score':'Adjusted_OppTempo'))

RegSeasonData = inner_join(RegSeasonData, teamWinPercentage, by=c('Season','TeamID'))

RegSeasonDataGrouped = RegSeasonData %>% select(.,-c('Outcome')) %>% group_by(.,Season,TeamName,TeamID) %>% summarise(.,across(c('Score':'OppDRPercentage'),mean))
RegSeasonDataGrouped = inner_join(RegSeasonDataGrouped,teamWinPercentage,by=c('Season','TeamID'))

```

# Get dataset with adjusted stats
```{r}
RegSeasonDataAdjusted = test %>% select(.,-c('Score':'OppPF')) %>% select(.,-c('GameLoc':'OppDRPercentage'))

RegSeasonDataGroupedAdjusted = RegSeasonDataAdjusted %>% select(.,-c('Outcome')) %>% group_by(.,Season,TeamName,TeamID) %>% summarise(.,across(c('Adjusted_Score':'Adjusted_OppTempo'),mean))

RegSeasonDataGroupedAdjusted = inner_join(RegSeasonDataGroupedAdjusted,teamWinPercentage,by=c('Season','TeamID'))
```

# Datasets of adjusted and non-adjusted statistics
```{r}
RegSeasonDataGroupedAll = inner_join(RegSeasonDataGrouped,RegSeasonDataGroupedAdjusted,by=c('Season','TeamName','TeamID')) %>% select(.,-c(WinPercentage.y)) %>% rename(.,WinPercentage=WinPercentage.x)

RegSeasonDataAll = inner_join(RegSeasonData,RegSeasonDataAdjusted,by=c('Season','DayNum','TeamName','TeamID','Outcome','OppTeamID'))
```

#Home Court Advantage

```{r}

#Add Game Location to dataset
RegSeasonDataAll_GameLoc = RegSeasonDataAll %>% group_by(Season,GameLoc) %>% select_if(is.numeric) %>% summarise(.,across(c('Score':'Adjusted_OppTempo'),mean))

#Get season averages for each stat
RegSeasonAll_Averages = RegSeasonDataAll %>% group_by(Season) %>% select_if(is.numeric) %>% summarise(.,across(c('Score':'Adjusted_OppTempo'),mean))

#Join season averages with game location averages
RegSeasonDataAll_GameLoc = inner_join(RegSeasonDataAll_GameLoc,RegSeasonAll_Averages,by=c('Season'))

#Change column names
colnames(RegSeasonDataAll_GameLoc) = gsub(".x","",colnames(RegSeasonDataAll_GameLoc))
colnames(RegSeasonDataAll_GameLoc) = gsub(".y","_SeasonAvg",colnames(RegSeasonDataAll_GameLoc))

#split dataset to subtract later
split1 = RegSeasonDataAll_GameLoc %>% select(-contains('_SeasonAvg'))
split2 = RegSeasonDataAll_GameLoc %>% select(contains('_SeasonAvg'))

#subtract split datasets
GameLocData = (split1 %>% ungroup() %>% select(Score:Adjusted_OppTempo)) - (split2 %>% ungroup() %>% select(Score_SeasonAvg:Adjusted_OppTempo_SeasonAvg))

#concat both datasets
RegSeasonDataAll_GameLoc = RegSeasonDataAll_GameLoc %>% select(Season,GameLoc)
GameLocData = cbind.data.frame(RegSeasonDataAll_GameLoc,GameLocData)

#new dataset with game location adjustments
RegSeasonDataAllHCA = inner_join(RegSeasonDataAll, GameLocData, by=c('Season','GameLoc'))

#Change column names
colnames(RegSeasonDataAllHCA) = gsub(".x","",colnames(RegSeasonDataAllHCA))
colnames(RegSeasonDataAllHCA) = gsub(".y","_GameLocDiff",colnames(RegSeasonDataAllHCA))

#split dataset to subtract later
split1 = RegSeasonDataAllHCA %>% select(-contains('_GameLocDiff'))
split2 = RegSeasonDataAllHCA %>% select(contains('_GameLocDiff'))

#subtract split datasets
GameLocData = (split1 %>% ungroup() %>% select(-c('Outcome','GameLoc','TeamName')) %>% select(Score:Adjusted_OppTempo)) - (split2 %>% ungroup() %>% select(-c()) %>% select(Score_GameLocDiff:Adjusted_OppTempo_GameLocDiff))

RegSeasonDataAllHCA = RegSeasonDataAllHCA %>% relocate(TeamID,TeamName,OppTeamID,Outcome,.before=Score) %>% select(Season,TeamID,TeamName,OppTeamID,Outcome)

RegSeasonDataAllHCA = cbind.data.frame(RegSeasonDataAllHCA,GameLocData)

RegSeasonDataAll = RegSeasonDataAll %>% relocate(TeamID, TeamName, OppTeamID, GameLoc, Outcome, .before=DayNum)

RegSeasonDataAllHCA
RegSeasonDataAll

# Adj_OffDef = RegSeasonDataAll_GameLoc %>% group_by(Season,TeamID) %>% select_if(is.numeric) %>% summarise(.,across(c('Score':'Adjusted_OppTempo'),mean)) %>% select(.,Season,TeamID,Adjusted_OE,Adjusted_DE) %>% rename(.,Adjusted_OE_SeasonAvg=Adjusted_OE,Adjusted_DE_SeasonAvg=Adjusted_DE)

# RegSeasonDataAll_GameLocGrouped = RegSeasonDataAll_GameLoc %>% group_by(.,Season,TeamID,TeamName,GameLoc) %>% select_if(is.numeric) %>% summarise(.,across(c('Score':'Adjusted_OppTempo'),mean))
# 
# RegSeasonDataAll_GameLocGrouped = inner_join(RegSeasonDataAll_GameLocGrouped,Adj_OffDef,by=c('Season','TeamID'))
# 
# colnames(RegSeasonDataAll_GameLocGrouped)
# RegSeasonDataAll_GameLocGrouped = RegSeasonDataAll_GameLocGrouped %>% mutate(.,Adjusted_OE_HCA = (Adjusted_OE[which(GameLoc == 'H')[1]]) - (Adjusted_OE_SeasonAvg)) %>% mutate(.,Adjusted_OE_ACA = (Adjusted_OE[which(GameLoc == 'A')[1]]) - (Adjusted_OE_SeasonAvg)) %>% mutate(.,Adjusted_OE_HCA_Pct = ( ((Adjusted_OE[which(GameLoc == 'H')[1]]) - (Adjusted_OE_SeasonAvg))/(Adjusted_OE_SeasonAvg)) * 100) %>% mutate(.,Adjusted_OE_ACA_Pct = (((Adjusted_OE[which(GameLoc == 'A')[1]]) - (Adjusted_OE_SeasonAvg))/(Adjusted_OE_SeasonAvg)) * 100)
# 
# colnames(RegSeasonDataAll_GameLocGrouped)
# RegSeasonDataAll_GameLocGrouped %>% mutate(Final_Adjusted_OE = Adjusted_OE * Adjusted_OE_HCA_Pct)


```

# Calculating strength of schedule
```{r}

AdjEff = RegSeasonDataAll %>% group_by(.,Season,TeamID) %>% summarise(Adjusted_OE=mean(Adjusted_OE), Adjusted_DE=mean(Adjusted_DE), Adjusted_EM=mean(Adjusted_EM))

AdjEff = AdjEff %>% rename(.,OppSeasonAdjusted_OE = Adjusted_OE,OppSeasonAdjusted_DE = Adjusted_DE,OppSeasonAdjusted_EM = Adjusted_EM)

RegSeasonDataAll = inner_join(RegSeasonDataAll,AdjEff,by=c("Season" = "Season","OppTeamID" = "TeamID"))
```

# Coming up with new model using linear regression
```{r}
RegSeasonDataAllGrouped = RegSeasonDataAll  %>% group_by(.,Season,TeamName, TeamID) %>% summarise(across(c('Score':'OppSeasonAdjusted_EM'),mean)) %>% select(.,-c('Outcome'))

RegSeasonDataAll = RegSeasonDataAll %>% mutate(.,ScoreMargin=Score-OppScore)
RegSeasonDataAll = RegSeasonDataAll %>% mutate(.,PointTotal=Score+OppScore)

#Perform model only on the 2021 season data
RegSeasonDataAll_2021 = RegSeasonDataAll %>% filter(.,Season=='2021')
RegSeasonDataAllGrouped_2021 = RegSeasonDataAllGrouped %>% filter(.,Season=='2021')

#Remove OppTeamID column before joining to the each season game for 2021 dataset
RegSeasonDataAllGrouped_2021 = RegSeasonDataAllGrouped_2021 %>% select(.,-c('OppTeamID'))
seasonAvgJoinedToGames = inner_join(RegSeasonDataAll_2021, RegSeasonDataAllGrouped_2021, by=c('Season','TeamName','TeamID'))

#Do data cleaning
seasonAvgJoinedToGames = seasonAvgJoinedToGames %>% select(.,-contains('.x'))
colnames(seasonAvgJoinedToGames) = gsub(".y","_Team_SeasonAvg",colnames(seasonAvgJoinedToGames))

seasonAvgJoinedToGames = inner_join(seasonAvgJoinedToGames, test_withOppAdjEff_Grouped_2021, by=c('Season'='Season','OppTeamID'='TeamID'))

#Do data cleaning
seasonAvgJoinedToGames = seasonAvgJoinedToGames %>% rename(.,TeamName=TeamName.x,OppTeamName=TeamName.y)
colnames(seasonAvgJoinedToGames)[118:224]=paste0(colnames(seasonAvgJoinedToGames)[118:224],"_Opp_SeasonAvg")

#Add Expected Tempo to dataframe
seasonAvgJoinedToGames = seasonAvgJoinedToGames %>% mutate(.,ExpectedTempo=as.numeric(season_averages[,'Adjusted_Tempo']) + (Adjusted_Tempo_Team_SeasonAvg - as.numeric(season_averages[,'Adjusted_Tempo'])) + (Adjusted_Tempo_Opp_SeasonAvg - as.numeric(season_averages[,'Adjusted_Tempo'])))

#Subset Data / Get rid of non numeric columns
seasonAvgJoinedToGames_Sub = seasonAvgJoinedToGames[c(7,9:225)]
seasonAvgJoinedToGames_Sub = seasonAvgJoinedToGames_Sub[c(-110)]

#Fit linear regression model
lm_ScoreMargin = lm(ScoreMargin~.,data=seasonAvgJoinedToGames_Sub)
summary(lm_ScoreMargin)


########## Predict the point margin for two teams ################
team = RegSeasonDataAllGrouped_2021 %>% filter(.,TeamName=='Wisconsin')
opp = RegSeasonDataAllGrouped_2021 %>% filter(.,TeamName=='Colgate')


colnames(team)[4:111]=paste0(colnames(team)[4:111],"_Team_SeasonAvg")
colnames(opp)[4:111]=paste0(colnames(opp)[4:111],"_Opp_SeasonAvg")

team = team %>% ungroup() %>% select(.,-c(Season, TeamName, TeamID))
opp = opp %>% ungroup() %>% select(.,-c(Season, TeamName, TeamID))

dataToPredict = cbind(team,opp)

#Add expected tempo to dataframe row
dataToPredict = dataToPredict %>% mutate(.,ExpectedTempo=as.numeric(season_averages[,'Adjusted_Tempo']) + (Adjusted_Tempo_Team_SeasonAvg - as.numeric(season_averages[,'Adjusted_Tempo'])) + (Adjusted_Tempo_Opp_SeasonAvg - as.numeric(season_averages[,'Adjusted_Tempo'])))
predicted_margin = predict(lm_ScoreMargin, newdata=dataToPredict)

########## Predict the point total for two teams ################

#Subset Data / Get rid of non numeric columns
seasonAvgJoinedToGames_Sub = seasonAvgJoinedToGames[c(8,9:225)]
seasonAvgJoinedToGames_Sub = seasonAvgJoinedToGames_Sub[c(-110)]

#Fit linear regression model
lm_PointTotal = lm(PointTotal~.,data=seasonAvgJoinedToGames_Sub)
summary(lm_PointTotal)

team = RegSeasonDataAllGrouped_2021 %>% filter(.,TeamName=='Wisconsin')
opp = RegSeasonDataAllGrouped_2021 %>% filter(.,TeamName=='Colgate')

colnames(team)[4:111]=paste0(colnames(team)[4:111],"_Team_SeasonAvg")
colnames(opp)[4:111]=paste0(colnames(opp)[4:111],"_Opp_SeasonAvg")

team = team %>% ungroup() %>% select(.,-c(Season, TeamName, TeamID))
opp = opp %>% ungroup() %>% select(.,-c(Season, TeamName, TeamID))

dataToPredict = cbind(team,opp)

#Add expected tempo to dataframe row
dataToPredict = dataToPredict %>% mutate(.,ExpectedTempo=as.numeric(season_averages[,'Adjusted_Tempo']) + (Adjusted_Tempo_Team_SeasonAvg - as.numeric(season_averages[,'Adjusted_Tempo'])) + (Adjusted_Tempo_Opp_SeasonAvg - as.numeric(season_averages[,'Adjusted_Tempo'])))

predicted_pointTotal = predict(lm_PointTotal, newdata=dataToPredict)

paste((predicted_pointTotal/2) + (predicted_margin/2), (predicted_pointTotal/2) - (predicted_margin/2),sep=' - ')
```

# Model experimentation
```{r}
colnames(seasonAvgJoinedToGames)

```

# Calculate number of wins for each team in the tournament
```{r}
tourneyResults <- read.csv(paste0(relativePath,'MDataFiles_Stage2/MNCAATourneyDetailedResults.csv'))
tourneyTeams <- read.csv(paste0(relativePath,'MDataFiles_Stage2/MNCAATourneySeeds.csv'))

tourneyTeams= tourneyTeams %>% select(.,Season,TeamID) %>% filter(Season>=2003)
numTourneyWins = tourneyResults %>% group_by(.,Season,WTeamID) %>% summarise(NumberOfTournamentWins=n())
tourneyTeams = left_join(tourneyTeams, numTourneyWins, by=c("Season" = "Season", "TeamID" = "WTeamID"))
tourneyTeams = tourneyTeams %>% replace(is.na(.), 0)

RegSeasonTT = inner_join(RegSeasonData,tourneyTeams,by=c('Season','TeamID'))
RegSeasonTTGrouped = inner_join(RegSeasonDataGroupedAll,tourneyTeams,by=c('Season','TeamID'))

```


```{r}
#fileName = paste(path_out, 'RegSeasonDataMain.csv',sep = '')
#write.csv(RegSeasonDataMain,fileName)
```


```{r}
#fileName = paste(path_out, 'RegSeasonTT.csv',sep = '')
#write.csv(RegSeasonTT,fileName)
```

```{r}
colnames(RegSeasonTTGrouped) = gsub("Opp","Opponent_",colnames(RegSeasonTTGrouped))
colnames(RegSeasonTTGrouped) = gsub("OE","Offensive_Efficiency",colnames(RegSeasonTTGrouped))
colnames(RegSeasonTTGrouped) = gsub("DE","Defensive_Efficiency",colnames(RegSeasonTTGrouped))
colnames(RegSeasonTTGrouped) = gsub("EM","Efficiency_Margin",colnames(RegSeasonTTGrouped))
colnames(RegSeasonTTGrouped) = gsub("Percentage","_Percentage",colnames(RegSeasonTTGrouped))
fileName = paste(path_out, 'RegSeasonTTGrouped.csv',sep = '')
write.csv(RegSeasonTTGrouped,fileName)
```

```{r}
RegSeasonData = RegSeasonData %>% filter(.,Season > 2016 )
colnames(RegSeasonData) = gsub("Opp","Opponent_",colnames(RegSeasonData))
colnames(RegSeasonData) = gsub("OE","Offensive_Efficiency",colnames(RegSeasonData))
colnames(RegSeasonData) = gsub("DE","Defensive_Efficiency",colnames(RegSeasonData))
colnames(RegSeasonData) = gsub("EM","Efficiency_Margin",colnames(RegSeasonData))
colnames(RegSeasonData) = gsub("Percentage","_Percentage",colnames(RegSeasonData))
fileName = paste(path_out, 'RegSeasonData.csv',sep = '')
write.csv(RegSeasonData,fileName)
```

```{r}
#RegSeasonDataAdjusted
#fileName = paste(path_out, 'RegSeasonDataAdjusted.csv',sep = '')
#write.csv(RegSeasonDataAdjusted,fileName)
```


```{r}
colnames(RegSeasonDataGrouped) = gsub("Opp","Opponent_",colnames(RegSeasonDataGrouped))
colnames(RegSeasonDataGrouped) = gsub("OE","Offensive_Efficiency",colnames(RegSeasonDataGrouped))
colnames(RegSeasonDataGrouped) = gsub("DE","Defensive_Efficiency",colnames(RegSeasonDataGrouped))
colnames(RegSeasonDataGrouped) = gsub("EM","Efficiency_Margin",colnames(RegSeasonDataGrouped))
colnames(RegSeasonDataGrouped) = gsub("Percentage","_Percentage",colnames(RegSeasonDataGrouped))
fileName = paste(path_out, 'RegSeasonDataGrouped.csv',sep = '')
write.csv(RegSeasonDataGrouped,fileName)
```

```{r}
#fileName = paste(path_out, 'RegSeasonDataGroupedAdjusted.csv',sep = '')
#write.csv(RegSeasonDataGroupedAdjusted,fileName)
```

```{r}
colnames(RegSeasonDataGroupedAll) = gsub("Opp","Opponent_",colnames(RegSeasonDataGroupedAll))
colnames(RegSeasonDataGroupedAll) = gsub("OE","Offensive_Efficiency",colnames(RegSeasonDataGroupedAll))
colnames(RegSeasonDataGroupedAll) = gsub("DE","Defensive_Efficiency",colnames(RegSeasonDataGroupedAll))
colnames(RegSeasonDataGroupedAll) = gsub("EM","Efficiency_Margin",colnames(RegSeasonDataGroupedAll))
colnames(RegSeasonDataGroupedAll) = gsub("Percentage","_Percentage",colnames(RegSeasonDataGroupedAll))
fileName = paste(path_out, 'RegSeasonDataGroupedAll.csv',sep = '')
write.csv(RegSeasonDataGroupedAll,fileName)
```

```{r}
#fileName = paste(path_out, 'SeasonAverages.csv',sep = '')
#write.csv(season_averages,fileName)
```

```{r}
#fileName = paste(path_out, 'TeamAverages.csv',sep = '')
#write.csv(team_averages,fileName)
```






